### 2025-08-10 modern-portfolio-theory.qmd
20250801 - correlation matrix plots
20250809 - label functions

```{r}
library("quantmod")
library("fPortfolio")
library("NMOF")
library("GGally")
library("tidyverse")
```

@@

In @fig-correlationMatrixPlot, we depict a correlation matrix plot using the `ggcorr()` function of the `GGally` [@R-GGally] package.

@@

GGally::ggcorr(combined_points_wide %>% select(QB1:opp_DST))
```

@@

@@

We use the `fPortfolio` package [@R-fPortfolio] to determine the optimal portfolio.

@@

@@

The tangency portfolio is the portfolio with the highest Sharpe ratio—i.e., the highest ratio of return to risk.
In other words, it is the portfolio with the greatest risk-adjusted returns.

```{r}
# Find the tangency portfolio (portfolio with the highest Sharpe ratio)
tangencyPortfolio <- fPortfolio::tangencyPortfolio(
  data = returns_ts,
  spec = portfolioSpec)

# Extract optimal weights
tangencyPortfolio_optimalWeights <- fPortfolio::getWeights(tangencyPortfolio)
tangencyPortfolio_optimalWeights

@@

If the `maxreturnPortfolio()` function worked as expected, the following could should work for determining the portfolio with the maximum return at each of various risk levels.
However, there is a known bug in the `fportfolio` package [@R-fPortfolio] that the optimal portfolio (based on maximum returns) does not change when changing the target risk level, suggesting that it is not taking into account the target risk level specified by the user (see here: <https://stackoverflow.com/q/78784306/2029527>).

@@

Below, we use the `mvFrontier()` function from the `NMOF` package [@R-NMOF; @Gilli2019_packages; @Schumann20112024_packages] to compute the efficient frontier.

@@

In summary, fantasy football is similar to stock picking.
You are most likely to pick the best players if you go with the [wisdom of the crowd](#sec-wisdomOfTheCrowd) (e.g., average projections across projection sources) and diversify.
Players—like stocks—have expected returns and risk associated with them.
You can think of constructing your team like constructing a stock portfolio, where your goal is to gain the maximum expected returns for a given level of risk.
The greater risk you are willing to take, the greater the potential returns, but alse the lower the potential downside if the players underperform.
Thus, you may want your team to be composed of a diverse combination of some higher risk players and some lower risk players.
We demonstrate the [efficient frontier](#sec-efficientFrontier) as a valuable tool for identifying the combination of players with the maximum expected returns relative to their risk.
Most projections are public information, so you might wonder whether using crowd projections gains you anything because everybody else has access to public information.
However, this is also the case with stocks, and people still consistently perform best over time when they go with the market.
Nevertheless, crowd projections are not highly accurate.
And fantasy football is a game, so feel free to have fun and deviate from the crowd!
However, you may be just as (if not more) likely to be wrong by deviating from the crowd.

>>>>

```{r}
library("quantmod")
library("TTR")
library("fPortfolio")
library("NMOF")
library("GGally")
library("tidyverse")
```

@@

In @fig-correlationMatrixPlot, we depict a correlation matrix plot using the `GGally::ggcorr()` function of the `GGally` [@R-GGally] package.

@@

GGally::ggcorr(
  combined_points_wide %>% select(QB1:opp_DST),
  label = TRUE,
  label_size = 3)
```

@@

We download closing prices using the `quantmod::Cl()` function of the `quantmod` package [@R-quantmod].
We calculate returns from the closing prices using the `TTR::ROC()` function of the `TTR` package [@R-TTR].

@@

We use the `fPortfolio` package [@R-fPortfolio] to determine the optimal portfolio.
We specify the portfolio specifications using the `fPortfolio::portfolioSpec()` function.

@@

We determine the efficient frontier using the `fPortfolio::portfolioFrontier()` function.

@@

The tangency portfolio is the portfolio with the highest Sharpe ratio—i.e., the highest ratio of return to risk.
In other words, it is the portfolio with the greatest risk-adjusted returns.
We identify the tangency portfolio using the `fPortfolio::tangencyPortfolio()` function.

```{r}
# Find the tangency portfolio (portfolio with the highest Sharpe ratio)
tangencyPortfolio <- fPortfolio::tangencyPortfolio(
  data = returns_ts,
  spec = portfolioSpec)
```

We extract the optimal weights for each asset at the tangency portfolio using the `fPortfolio::getWeights()` function.

```{r}
# Extract optimal weights
tangencyPortfolio_optimalWeights <- fPortfolio::getWeights(tangencyPortfolio)
tangencyPortfolio_optimalWeights

@@

If the `fPortfolio::maxreturnPortfolio()` function worked as expected, the following could should work for determining the portfolio with the maximum return at each of various risk levels.
However, there is a known bug in the `fportfolio` package [@R-fPortfolio] that the optimal portfolio (based on maximum returns) does not change when changing the target risk level, suggesting that it is not taking into account the target risk level specified by the user (see here: <https://stackoverflow.com/q/78784306/2029527>).

@@

Below, we use the `NMOF::mvFrontier()` function from the `NMOF` package [@R-NMOF; @Gilli2019_packages; @Schumann20112024_packages] to compute the efficient frontier.

@@

In summary, fantasy football is similar to stock picking.
You are most likely to pick the best players if you go with the [wisdom of the crowd](#sec-wisdomOfTheCrowd) (e.g., average projections across projection sources) and diversify.
Players—like stocks—have expected returns and risk associated with them.
You can think of constructing your team like constructing a stock portfolio, where your goal is to gain the maximum expected returns for a given level of risk.
The greater risk you are willing to take, the greater the potential returns, but also the lower the potential downside if the players underperform.
Thus, you may want your team to be composed of a diverse combination of some higher risk players and some lower risk players.
We demonstrate the [efficient frontier](#sec-efficientFrontier) as a valuable tool for identifying the combination of players with the maximum expected returns relative to their risk.
Most projections are public information, so you might wonder whether using crowd projections gains you anything because everybody else has access to public information.
However, this is also the case with stocks, and people still consistently perform best over time when they go with the market.
Nevertheless, crowd projections are not highly accurate.
And fantasy football is a game, so feel free to have fun and deviate from the crowd!
However, you may be just as (if not more) likely to be wrong by deviating from the crowd.

### 2025-08-10 multiple-regression.qmd
20250806 - robust load
20250809 - label functions

We fit a linear regression model using the `lm()` function.

@@

We fit a logistic regression model using the `glm()` function and specifying `family = binomial()`.
To calculate the model $R^2$, we use the `r2()` function of the `performance` package [@R-performance; @Ludecke2021_packages].

@@

We fit an ordinal regression model using the `clm()` function of the `ordinal` package [@R-ordinal].

@@

We fit a Poisson regression model using the `glm()` function and specifying `family = poisson()`.

@@

We fit a negative binomial regression model using the `glm.nb()` function of the `MASS` package [@R-MASS].

@@

We fit a (Bayesian) zero-one-inflated beta regression model (to allow for zeros and ones) using the `brm()` function of the `brms` package [@R-brms] and specifying `family = zero_one_inflated_beta()`.

@@

load(url("https://osf.io/download/fe37j/"))
```

@@

If we want to obtain standardized regression coefficients, we can use the `standardize_parameters()` function of the `effectsize` package [@R-effectsize; @BenShachar2020_packages].

@@

Or, we can standardize the [outcome variable](#sec-correlationalStudy) and each [predictor variable](#sec-correlationalStudy) using the `scale()` function:

@@

We can perform a dominance analysis to evaluate the relative importance of the predictors in the regression model.
To perform the dominance analysis, we use the `dominance_analysis()` function of the `parameters` package [@R-parameters; @Luedecke2020_packages], which leverages the `domir` package [@R-domir].

@@

To visualize the regression coefficients, we can use the `tidy()` function of the `broom` package [@R-broom], as in Figures [-@fig-regressionCoefficients] and [-@fig-regressionCoefficientsStandardized].

@@

An important consideration in multiple regression is how missing data are handled.
Multiple regression in `R` using the `lm()` function applies listwise deletion.
*Listwise deletion* (also called complete case analysis) removes any row (in the data file) from analysis that has a missing value on the [outcome variable](#sec-correlationalStudy) or any of the [predictor variables](#sec-correlationalStudy).
Removing all rows from analysis that have any missingness in the model variables can be a problem because missingness is often not completely at random—missingness often occurs systematically (i.e., for a reason).
For instance, participants may be less likely to have data for all variables if they are from a lower socioeconomic status background and do not have the time to participate in all study procedures.
Thus, applying listwise deletion, we might systematically exclude participants from lower socioeconomic status backgrounds (or other groups), which could lead to less generalizable inferences.

@@

To address outliers, there are various approaches to robust regression.
One approach is to use an MM-type estimator, such as is used in the `lmrob()` and `glmrob()` functions of the `robustbase` package [@R-robustbase; @Todorov2009_packages].

@@

We set up the bootstrap folds using the `bootstraps()` function of the `rsample` package [@R-rsample].

>>>>

We fit a linear regression model using the `stats::lm()` function.

@@

We fit a logistic regression model using the `stats::glm()` function and specifying `family = binomial()`.
To calculate the model $R^2$, we use the `performance::r2()` function of the `performance` package [@R-performance; @Ludecke2021_packages].

@@

We fit an ordinal regression model using the `ordinal::clm()` function of the `ordinal` package [@R-ordinal].

@@

We fit a Poisson regression model using the `stats::glm()` function and specifying `family = poisson()`.

@@

We fit a negative binomial regression model using the `MASS::glm.nb()` function of the `MASS` package [@R-MASS].

@@

We fit a (Bayesian) zero-one-inflated beta regression model (to allow for zeros and ones) using the `brms::brm()` function of the `brms` package [@R-brms] and specifying `family = zero_one_inflated_beta()`.

@@

petersenlab::robust_load("https://osf.io/download/fe37j/") #load(url("https://osf.io/download/fe37j/"))
```

@@

If we want to obtain standardized regression coefficients, we can use the `effectsize::standardize_parameters()` function of the `effectsize` package [@R-effectsize; @BenShachar2020_packages].

@@

Or, we can standardize the [outcome variable](#sec-correlationalStudy) and each [predictor variable](#sec-correlationalStudy) using the `base::scale()` function:

@@

We can perform a dominance analysis to evaluate the relative importance of the predictors in the regression model.
To perform the dominance analysis, we use the `parameters::dominance_analysis()` function of the `parameters` package [@R-parameters; @Luedecke2020_packages], which leverages the `domir` package [@R-domir].

@@

To visualize the regression coefficients, we can use the `broom::tidy()` function of the `broom` package [@R-broom], as in Figures [-@fig-regressionCoefficients] and [-@fig-regressionCoefficientsStandardized].

@@

An important consideration in multiple regression is how missing data are handled.
Multiple regression in `R` using the `stats::lm()` function applies listwise deletion.
*Listwise deletion* (also called complete case analysis) removes any row (in the data file) from analysis that has a missing value on the [outcome variable](#sec-correlationalStudy) or any of the [predictor variables](#sec-correlationalStudy).
Removing all rows from analysis that have any missingness in the model variables can be a problem because missingness is often not completely at random—missingness often occurs systematically (i.e., for a reason).
For instance, participants may be less likely to have data for all variables if they are from a lower socioeconomic status background and do not have the time to participate in all study procedures.
Thus, applying listwise deletion, we might systematically exclude participants from lower socioeconomic status backgrounds (or other groups), which could lead to less generalizable inferences.

@@

To address outliers, there are various approaches to robust regression.
One approach is to use an MM-type estimator, such as is used in the `robustbase::lmrob()` and `robustbase::glmrob()` functions of the `robustbase` package [@R-robustbase; @Todorov2009_packages].

@@

We set up the bootstrap folds using the `rsample::bootstraps()` function of the `rsample` package [@R-rsample].

### 2025-08-10 pca.qmd
20250809 - label functions

We generated the scree plot in @fig-pcaScreePlot1 using the `fa.parallel()` function of the `psych` package [@R-psych].

@@

We generated the scree plot in @fig-pcaScreePlot2 using the `nScree()` and `plotnScree()` functions of the `nFactors` package [@R-nFactors].

@@

We generated the very simple structure (VSS) plots in Figures [-@fig-pcaVSSPlot1] and [-@fig-pcaVSSPlot2] using the `vss()` and `nfactors()` functions of the `psych` package [@R-psych].
The optimal number of components based on the VSS criterion is three components.

>>>>

We generated the scree plot in @fig-pcaScreePlot1 using the `psych::fa.parallel()` function of the `psych` package [@R-psych].

@@

We generated the scree plot in @fig-pcaScreePlot2 using the `nFactors::nScree()` and `nFactors::plotnScree()` functions of the `nFactors` package [@R-nFactors].

@@

We generated the very simple structure (VSS) plots in Figures [-@fig-pcaVSSPlot1] and [-@fig-pcaVSSPlot2] using the `psych::vss()` and `psych::nfactors()` functions of the `psych` package [@R-psych].
The optimal number of components based on the VSS criterion is three components.

### 2025-08-10 simulation.qmd
20250809 - label functions

For performing the bootstrapping, we leverage the `data.table` [@R-data.table], `future` [@R-future], and `future.apply` [@R-future.apply] packages for speed (by using parallel processing) and memory efficiency.
We use the `progressr` [@R-progressr] package to create a progress bar.

@@

@@

@@

For each player, we identify the optimal distribution as either a normal distribution , or as a skew-normal distribution.
The normal distribution was fit using the `fitdist()` function of the `fitdistrplus` package [@R-fitdistrplus; @DelignetteMuller2015_packages].
The skew-normal distribution was fit using the `selm()` function of the `sn` package [@R-sn; @Azzalini2023_packages].

@@

Now, we can run the model.

>>>>

For performing the bootstrapping, we leverage the `data.table` [@R-data.table] (`data.table::as.data.table()`; `data.table::data.table()`; `data.table::rbindlist()`), `future` [@R-future] (`future::plan()`; `future::multisession()`), and `future.apply` [@R-future.apply] (`future.apply::future_lapply()`) packages for speed (by using parallel processing) and memory efficiency.
We use the `progressr` [@R-progressr] package (`progressr::handlers()`; `progressr::with_progress()`; `progressr::progressor()`) to create a progress bar.

@@

`data.table::setnames()`

@@

`base::lapply()`

@@

For each player, we identify the optimal distribution as either a normal distribution , or as a skew-normal distribution.
The normal distribution was fit using the `fitdistrplus::fitdist()` function of the `fitdistrplus` package [@R-fitdistrplus; @DelignetteMuller2015_packages].
The skew-normal distribution was fit using the `sn::selm()` function of the `sn` package [@R-sn; @Azzalini2023_packages].

@@

Now, we can run the model using the `SimDesign::runSimulation()` function.

### 2025-08-10 time-series-analysis.qmd
20250806 - robust load
20250809 - label functions

```{r}
load(url("https://osf.io/download/q6rjf/")) # Bayesian model object
```

@@

The autocorrelation function (ACF) plot depicts the autocorrelation of scores as a function of the length of the lag.
Significant autocorrelation is detected when the autocorrelation exceeds the dashed blue lines, as is depicted in @fig-timeSeriesACFtomBrady.

@@

Using the `forecast` package [@R-forecast; @Hyndman2008_packages]:

@@

@@

@@

>>>>

```{r}
petersenlab::robust_load("https://osf.io/download/q6rjf/") #load(url("https://osf.io/download/q6rjf/")) # Bayesian model object
```

@@

The autocorrelation function (ACF) plot depicts the autocorrelation of scores as a function of the length of the lag.
We can generate an ACF plot using the `stats::acf()` function.
Significant autocorrelation is detected when the autocorrelation exceeds the dashed blue lines, as is depicted in @fig-timeSeriesACFtomBrady.

@@

We can fit an ARIMA model using the `forecast::auto.arima()` function of the `forecast` package [@R-forecast; @Hyndman2008_packages]:

@@

We can generate a plot of an ARIMA model using the `stats::arima()` function.

@@

We can generate model forecasts from the ARIMA model using the `forecast::forecast()` function.

@@

We can plot the model forecasts using the `forecast::autoplot()` function.
